---
title: "Report1"
author: "David Kyle Breault"
date: "9/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
source("CleanData.R")
```

## Loading and Cleaning Data

Obtained multiple datasets of the same variables in order to have a larger collection of data, which were parsed and joined together. The final dataset (zoo.final) had two additionally mutated columns (see `Dataset Information`). There we no `NA` values in any of the joined, or final, dataset.

For parsing, every entry (with the exception of animal_name which was a character) was entered as a double. However, all the entries listed as doubles, not counting `legs` or `class_type` used the 1 and 0 to represent TRUE and FALSE. For this reason, all of those entries were parsed manually to be a logical, since that is what the 1 and 0's were for Boolean data. `legs` was left as a double, and `animal_name` left as a character; however, `class_type` was parsed as a character, due to the fact that each number in the class_type column represented a specific animal (see below).

For class_types, the names are numeric but represent character values (species type) so adding a row representing `species_type` was helpful. This was done by using the class_type names (what species type represents what number) as specified in the data summary. 1 being mammal, 2 bird, 3 reptile, 4 fish, 5 amphibian, 6 insect, 7 non-insect invertebrate (Note: 6 is bug in the data summary, and 7 is invertebrate in the data summary. However, all bugs are insects, but not all insects are bugs, and for class_type 6, insect is more accurate given the species in the data set (as some of the species given are insects but not bugs)). For 7, invertebrate is too broad of a term, as that includes insects, so this name designation was given non-insect invertebrates as a result).

The `venomous` column name was changed to `venomous.poisonous`, as there is a distinctive difference between a venomous species and a poisonous species. When exploring data on amphbians, it was noted that some species listed were venomous instead of poisonous. Evidence of this is ra-flecha-azul (which is portugese), also known as the blue arrow frog, which is poisionous. Sapo-cururu, portugese for Rhinella icterica, is also poisonous, but listed as venomous. This issue could be prevelant in other areas of the dataset. For this reason, the column name was changed to `venomous.poisonous`.

There are also errors in the `legs` column; however, no changes were made to that column at this time (see errors in `Exploratory Analysis`).

## Dataset information

Three datasets were utalized in exploratory analysis:

zoo.csv: 18 columns, 101 rows.

What it measures: Dataset containing information on a variety of animal species that could be helpful in machine learning. The data for this project will not be used for machine learning, but will instead be interpreted by a person to look for relationships. This data is used for classification, in other words, based off of what anatomical features this species has, how should it be listed taxonomically?

How was it collected: Dataset was collected from a (unspecified) zoo. Collected by UCI Machine Learning (https://www.kaggle.com/uciml/zoo-animal-classification/metadata)

What it contains: 18 columns, with hair, feathers, eggs, milk, airborne, aquatic, predator, toothed, backbone, breahtes, venomous, fins, tail, domestic, and catsize being boolean. animal_name is unique for each row, and class_type is numeric, but represents a specific character (species type). legs are numeric, and are in values of 0,2,4,6,8.

What columns were used: venomous.posionous, animal_name, legs,and aquatic

zoo2.csv: 18 columns, 43 rows

What does it measure: an extension of the original zoo.csv dataset with more species that are not mammals or birds in order to help make a more diverse dataset to determine what pararameters are important for classifying species.

How ws it collected: unspecified. Collected by Rodrigo Hjort (https://www.kaggle.com/agajorte/zoo-animals-extended-dataset)

What columns were used: venomous.posionous, animal_name, legs,and aquatic

zoo3.csv: 18 columns, 70 rows

What does it measure: an extension of the original zoo.csv dataset with more species being mainly animals from brazil that are in danger of extinction

How was it collected: São Paulo Zoo. Collected by Rodrigo Hjort (https://www.kaggle.com/agajorte/zoo-animals-extended-dataset)

What columns were used: venomous.posionous, animal_name, legs,and aquatic

zoo.final: 20 columns, 214 rows

What does it measure: zoo.final is zoo.csv, zoo2.csv, and zoo3.csv joined together. This dataset measures the same paramaters as each individual dataset. Two mutated columns were added, being species_type and data origin.

What columns were used: venomous.posionous, animal_name, legs,and aquatic (original dataset) species_type (mutated extra column)


## Exploratory Analysis

What are the trends with venomous species?

```{r}
zoo.final %>%
  group_by(species_type) %>%
  summarize(Percent_venomous = mean(venomous.poisonous)) %>%
  ggplot()+
  geom_col(aes(y=Percent_venomous,x=species_type,fill=species_type))+
  labs(title="Species that are Venomous",
       x="Class Type",
       y="Percent of Venomous Species", fill = "Species")
```


For the summarize function, the mean of venomous species was taken. This is because taking the average of a logical will result in a percent. Because there are a different amount of entries for each `species_type`(ie: how many total mammals there are in the data set vs fish) taking the percent will give a more accurate count of the data to understand relationships.

Most of the species that are venomous in this dataset are non insect invertebrates. This number is most likley do to the prevalence of arachnids which are commonly venomous.The bar for amphibians is noted by the poisonous species. For reptiles what species be driving up the value of venomous species? Reptiles are made up of species that are snakes (0 legs) and non-snakes (4 legs). Are non-snakes or snakes driving up the distribution of venomous species.

```{r}
zoo.final%>%
  filter(species_type=="reptile", animal_name!="slowworm")%>%
  select(animal_name,species_type,venomous.poisonous,legs)%>%
  mutate(snake=
           case_when(
             legs=="4"~"non-snake",
             legs=="0"~"snake"
           ))%>%
  ggplot()+
geom_bar(mapping=aes(x=snake,fill=venomous.poisonous))+
  labs(title="Species that are Venomous",
       x="Snake or not a Snake",
       y="Number of Species", fill = "Venomous")

## One species incldued, being slowworm is a legless lizard, so it was filtered out for this analysis
```

How is a single variable (legs) represented in the dataset?

```{r}
zoo.final%>%
  summarise(avg_legs=mean(legs))
```

The summarise function was used to create summary statistics for the average number of legs in the whole dataset. Most terrestrial vertebrate groups have four legs, being mammals, reptiles, and amphibians with the exception of 2 for birds. Most of the dataset falls into a terrestrial vertebrate group, so waht could be bring down this average

```{r}
zoo.final%>%
  group_by(species_type)%>%
  summarise(avg_legs=mean(legs))%>%
  ggplot(aes(x=species_type,y=avg_legs))+
    geom_col(aes(fill=species_type))+ labs(title="Average Number of Legs",
       x="Class Type",
       y="Average Leg Number", fill = "Species")

```

All parts of the distribution make sense except for mammals and  amphbians which deviate from the expected (4 legs). Birds have 2 legs (expected), insect have 6 (expected), noninsect invertebrates is about 3 (expected, arachnids (8 legs) and other invertebrates like molluscs (0 legs)) and reptile with a little over 2 (expected, snakes bring down the average)

Take a look at amphbians first:

```{r}
zoo.final%>%
  filter(species_type=="amphibian")%>%
  select(animal_name,legs)
```

For amphibians, the leg average is being brought down by a siren, which was correctly listed as having two legs.

For mammals however, unless an aquatic species (such as dolphins or whales) are in the dataset, a leg value less than 4 is not expected.

```{r}
## Let's see if aquatic species bring down the average

zoo.final%>%
  filter(aquatic==TRUE, species_type=="mammal")%>%
  select(aquatic,species_type,animal_name, legs)

```

For mammals, the average leg number is brought down by these aquatic species. However, some are reported as having legs when they do not, such as the sealion and peixe-boi (manatee). While the sealion has functional legs (can use flippers as legs on land) they lack true leg anatomy. In addition, manatees are solely aquatic species, meaning they lack even functional legs.

## Conclusion

This dataset is constrained by the sample size. Even with the added data entries signifigant anaylsis is limited. An even larger data set with a randomization of species (not constrained to ones at a zoo) may be better for machine learning and classification purposes. In addition, some of the information is not classified correctly (such as venomous in amphibians,or inaccurate leg count, which may skew the ability of proper machine learning.) 

I believe that the variables themselves are diagnostic for classifying species and will be ample for machine learning. Seperating out venomous and poisonous species should occur. However, the datasets need to be larger in order to have a more represenative sample. The sample is too small to make any signifigant analysis. Further questions could explore the impact of `venomous` on other variables besides species type, such as aquatic, legs, and catsize. Looking at the relationship between domestication and catsize may also prove to be an interesting analysis. Further studies should continue to check the accuracy of the data and note any inconsistencies that appear.
